---
title: "Lymph_Prostate"
output:
  html_document: default
  word_document: default
  pdf_document: default
---

---
title: "From scratch"
author: "Aaron Patterson"
date: "25/05/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
library("TwoSampleMR")
library("MendelianRandomization")
library("R.utils")
library("MendelianRandomization")
```

##Load exposure data - In this case Lymphocyte GWAS summary data from Astle et al available on GWAS catalog


```{r }
lymphro.path = "/rdsgpfs/general/user/arp119/home/data/"
Lymphocyte.data = read.csv(paste(lymphro.path,"27863252-GCST004627-EFO_0004587.h.tsv.gz", sep=""), stringsAsFactors=FALSE, sep="\t")
```
##Subset Lymphocyte p-values <5x10^-8. For IV assumption 1.
```{r}
Lymphocyte.iv = Lymphocyte.data[Lymphocyte.data$p_value <5*10^-8,]

```

###Load outcome data from PRACTICAL Consortium. Summary meta-analysis data for 12 studies in European ancestry

```{r}
prostate.path = "/rdsgpfs/general/user/arp119/home/data/"
prostate.data = read.csv(paste(prostate.path,"meta_v3_onco_euro_overall_ChrAll_1_release.txt",
sep=""),stringsAsFactors=FALSE, sep="\t")
```

## significant p-values 5x10^-8 for outcome for valid outcome variable
```{r}

prostate.data = prostate.data[prostate.data$Pvalue<5*10^-8,]
```

##recode rsid for both exposure and outcome
```{r}
colnames(prostate.data)[3]="rs"
colnames(Lymphocyte.iv)[2]="rs"

```

##merge datasets by SNP

```{r}
data.merge=merge(Lymphocyte.iv, prostate.data, by="rs")

```
##recode merged esposure data to keep track of variables
```{r}
colnames(data.merge)[c(6,5)] = c("lymph_effectallele","lymph_noneffectallele")
colnames(data.merge)[c(7,22,23)] = c("lymph_beta","lymph_se","lymph_pval")
colnames(data.merge)[11] = c("lymph_eaf")
```

##recode merged outcome data to keep track of variables
```{r}
colnames(data.merge)[c(35,36,38)] = c("prost_effectallele","prost_noneffectallele","prost_eaf_not_aligned")
colnames(data.merge)[c(41,42,43)] = c("prost_beta_not_aligned","prost_se","prost_pval")
```

##Convert lower to upper case alleles
```{r}
table(data.merge$lymph_effectallele == toupper(data.merge$prost_effectallele))

table(data.merge$lymph_effectallele == toupper(data.merge$prost_noneffectallele))
```


##Following code for inconsistent values. Not applicable in this case
```{r}

###inconsistent=which(data.merge$lympho_effectallele) != data.merge$prost_effectallele& data.merge$lympho_effectallele) != data.merge$prost_noneffectallele)
###data.merge[inconsistent,]
###data.merge = data.merge[-inconsistent,]
```
##Harmonization of outcome variable
```{r}
data.merge$prost_eaf = ifelse(tolower(data.merge$lymph_effectallele) == data.merge$prost_effectallele,
data.merge$prost_eaf_not_aligned, 1-data.merge$prost_eaf_not_aligned)
```

##Pruning
```{r}
colnames(data.merge)[1]="SNP"
colnames(data.merge)[23]="pval.exposure"
```
##Clumping 0.001
```{r}
data.clump = clump_data(data.merge)
dim(data.clump)
```
##Variables for MR
```{r}

rs = data.clump$SNP
lymph_beta = data.clump$lymph_beta
lymph_se = data.clump$lymph_se
prost_beta = data.clump$prost_beta
prost_se = data.clump$prost_se
```
##Performing MR with IVW output
```{r, echo=FALSE}
mr.input = mr_input(bx = lymph_beta, bxse = lymph_se, by = prost_beta, byse = prost_se, exposure = "lymph", outcome = "Prostate", snps = rs)
mr_ivw(mr.input)
```
##Weighted IVW
```{r, echo=FALSE}
summary(lm(prost_beta~lymph_beta -1, weights=prost_se^(-2)))
```
## Plot
```{r, echo=FALSE}
mr_plot(mr.input, interactive=FALSE)
```
## Sensitivity analysis
```{r, echo=FALSE}
mr_allmethods(mr.input)
```
##Plot of all methods
```{r, echo=FALSE}
mr_plot(mr_allmethods(mr.input, method = "main"))
```

